
%%% lorem.tex --- 
%% 
%% Filename: lorem.tex
%% Description: 
%% Author: Ola Leifler
%% Maintainer: 
%% Created: Wed Nov 10 09:59:23 2010 (CET)
%% Version: $Id$
%% Version: 
%% Last-Updated: Wed Nov 10 09:59:47 2010 (CET)
%%           By: Ola Leifler
%%     Update #: 2
%% URL: 
%% Keywords: 
%% Compatibility: 
%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 
%%% Commentary: 
%% 
%% 
%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 
%%% Change log:
%% 
%% 
%% RCS $Log$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 
%%% Code:

\chapter{Conclusion}
\label{cha:conclusion}

In this thesis, we have shown that LogClustering is the best performing offline algorithm based on all three evaluation metrics (precision, recall and F1) on the HDFS and BGL dataset. As LogClustering uses a knowledge base, with stored sequences from each cluster which it compares new sequences to, it is highly suitable for system log files analysis. Additionally, LogClustering is the only technique out of the three offline which trains on normal data, making it easier to separate anomalous data when arriving in the testing phase. Therefore, LogClustering performs better than Isolation Forest and PCA in an offline setting. 

This study compared two online anomaly detection techniques, Half-Space Trees and an online implementation of One-Class SVM, each implemented on two datasets, HDFS and BGL. In this case, the results do not indicate a best performing algorithm, as Half-Space Trees performed slightly better on the HDFS dataset, and One-Class SVM performed better on the BGL dataset. HST is however particularly effective when applied to datasets with few anomalous data points, which is proven since it performs better on HDFS which contains less anomalies than BGL. Further, One-Class SVM performs better on dataset with more complex, deviated anomalies than HST, due to it's feature space boundary. With this said, HST is the more suitable algorithm when the proportion of anomalies within the dataset is relatively low, and easy to isolate. While One-Class SVM is better suited on scenarios with a more complex, deviated, and higher percentage of anomalies.

The results for the offline and online anomaly detection techniques differed severely. Offline performed significantly better in regards to all three evaluation metrics. The average F1-score was $50$ percentage points higher for the offline methods on the HDFS data and $57$ percentage points higher on the BGL dataset. There are several reasons behind this, the biggest being that offline methods trained on the full dataset and online methods only trained on $80$\%. Also, since offline algorithms have access to the data the evaluation is performed on beforehand which has a huge impact. Lastly, a reason for the big result gap is that the online methods was much slower and our experimental setup couldn't handle tweaking the parameters of the algorithms as we did with the offline techniques.  

The conclusions made us achieve our aim of comparing various anomaly detection techniques in different non-encrypted log datasets. We found \revrev{what algorithm was the best}{which algorithm performed best
} on two different datasets and compared the two different use cases, offline and online, which also was a part of the aim for this study.

\section{Future work}
As this study investigates online and offline anomaly detection on log data, evaluating the three metrics precision, recall and F1-score, it leaves space for future work. The need \revrev{of}{for} online anomaly detection is huge, and there are more anomaly \revrev{detections}{detection} techniques to be investigated. Further, finding anomalies \revrev{are}{is} important, and the three metrics are fully relevant. However in an online setting additional factors as CPU usage and running time \revrev{comes of importance}{is of importance}. This thesis provides a good foundation to further investigate these metrics, with the potential to further strengthen security on log data. \revrev{}{Also, to better judge the potential of the online algorithms, future work should focus on tuning their parameters. As improved tuning may lead to more fair comparisons and reveal whether the current performance gap between offline and online models is due to limitations in the models or simply a lack of optimization.}
Because of a limited time scope, this thesis used one general method with one log parsing technique, two different feature extractors and five anomaly detection techniques applied on two different datasets. It would be interesting for future research to compare other factors than anomaly detection, such as general methods, log parsing and more feature extractors to see how preprocessing data affects the overall result. Additionally, studying more datasets, with different structures would help \revrev{}{in} finding the best solutions for different situations. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% lorem.tex ends here

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "demothesis"
%%% End: 
